
title: "Lecture Guide"
author: "[your name]"
date: "11/01/2021"
output: 
  xaringan::moon_reader:
    seal: true 
    css: custom_2.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(tidyverse)
```

# Outline 

1. Recap: RMD, pipes, `select`, `filter` 
2. `summarize` +
3. Exercises 

## RMD

- Final draft/determine whether or not you need to include the code chunk in the output 

- One code chunk per output 

### This is what we want 

First, we also look at the total population across country-years: 

```{r, eval = F}
gapminder %>%  
  select(pop) %>% 
  summary()
```

Then, we look at the summary statistics of life expectancy across all country-years:
```{r, eval = F}
gapminder %>%  
  select(lifeExp) %>% 
  summary()
```

### We do not want 

I can't write about bother things in a smooth way... Maybe they output separately, but I can't write about them ...

```{r, eval = T}
gapminder %>%  
  select(pop) %>% 
  summary()

gapminder %>%  
  select(lifeExp) %>% 
  summary()
```

... unless I do it in this awkward sandwich style.


## Pipes 

- Read %>% as "then"

- **It fills in the first argument of your function from the tibble previously created**

- It lets us perform multiple operations at once, without creating an object for each one of them.


## How it works

Piped syntax: 
```{r eval = F}
gapminder %>% 
  select(country, year, lifeExp)
```

- This output is now a tibble, when we add a pipe, we are saying take this modified tibble (i.e. with only 3 variables/columns) and use this as our data source

- When our next `dplyr` function like `filter` looks for the first argument, e.g. 'data', it will use that modified tibble

## Filter

- `select` selects variables/columns

- `filter` filters out rows/observations/respondents

- **REQUIRES LOGICAL STATEMENT**: IF the statement about the data is TRUE, THEN it will give you those rows

## Logical Notations 

Symbol |Meaning 
-|-
==|*is exactly equal to*
!=|*is not equal to*
&|*and*
\||*or*
>=|*greater than or equal to*
<=|*less than or equal to*
>|*greater than*
<|*less than*
itTRUE(x)|*tests whether x is T*

# Outline 

1. Recap: Pipes, `select`, `filter`, RMD 
2. **`summarize`+**
3. Exercises

## `summarize`

- Recall: `summarize` condenses into the number of summary stats you are asking for

```{r}
gapminder %>% 
  summarize(blah_chicken = mean(pop))
```

## `summarize` by groups

- use `group_by`

```{r}

```


- Exercise 1: Before running this code, figure out how many rows and columns you expect to see (hint: use tables) 


- Exercise 2: What about if you grouped by *both* country and *year* 


## `summarize` multiple 

```{r}

```

- Exercise 3: Before running this code, figure out how many rows and columns you expect to see.


# Exercises 

Use `iris` data and R scripts + RMD

1. Find the descriptive stats on sepal length and write a sentence about it in the RMD 
2. Find the average petal length and width per each species 
3. Do the same but only for those flowers with a petal length under 5.0 and petal width above 1
4. Print the result of number 3 into a nice table in RMD and talk through the results 


## Bonus 

5. Bonus (easy): what does `mutate` do?
6. Bonus (hard): add a column call `petal_ratio` that is the ratio of petal length and petal width
7. Bonus (extreme): add two columns, one called `petal_ratio` and the other called `sepal_ratio` while using `mutate` once (hint: see how we summarized multiple)


